
::: {.callout-caution collapse="true"}
## Practice exercises: Advanced `create_template()` use cases

1. Create a stock assessment report template tailored for the AFSC (a SAFE report).

2. Create a stock assessment report template with two main sections: an executive summary and an introduction (the following sections will be added automatically: acknowledgments, references, tables, figures).

3. Whoops! You forgot to add your species name, and now you don't have a cover image. Run `create_template()` with the argument `rerender_skeleton = TRUE`, and add your species name (common and scientific names) as two additional arguments: "Petrale sole" (*Eopsetta jordani*).

<!-- 
Add questions to practice adding new sections and calling a previous report when features are ready.
2. Create a stock assessment report template with two new sections. "" should be placed before the introduction and "" should be placed after the acknowledgments. -->

:::

::: {.callout-caution collapse="true"}
## **Answers** for Practice exercises: Advanced `create_template()` use cases

1. Create a stock assessment report template tailored for the AFSC (a SAFE report).

**Answer**:

```{r eval=FALSE}
asar::create_template(
  type = "safe"
)
```

<!-- asar::create_template( -->
<!--   # custom = TRUE, -->
<!--   # new_section = c("Changes to the Model", "Sec3"), -->
<!--   # section_location = c("before-introduction", "after-acknowledgments") -->
<!-- ) -->

2. Create a stock assessment report template with two main sections: an executive summary and an introduction (the following sections will be added automatically: acknowledgments, references, tables, figures).

**Answer**:

```{r eval=FALSE}
asar::create_template(
    authors = c("Samantha Schiano" = "OST"),
    custom = TRUE,
    custom_sections = c("executive_summary", "introduction")
)
```

3. Whoops! You forgot to add your species name, and now you don't have a cover image. Run `create_template()` with the argument `rerender_skeleton = TRUE`, and add your species name (common and scientific names) as two additional arguments: "Petrale sole" (*Eopsetta jordani*).

**Answer**:

```{r eval=FALSE}
asar::create_template(
  rerender_skeleton = TRUE,
  species = "Petrale sole",
  spp_latin = "Eopsetta jordani",
  file_dir = file.path("report")
)
```

:::
