
::: {.callout-caution collapse="true"}
## Practice exercises: Creating `stockplotr` figures & tables

For the following, use `stockplotr::example_data` as the input data.

1. Create an abundance-at-age plot. Export it as an rda to your working directory.
2. Create a recruitment time series plot. Change the `unit_label` and `scale_amount` so that the y axis shows recruitment in thousands of fish. Set `module` = "TIME_SERIES".
3. Create a fishing mortality plot. Ensure it does not group by fleet. Set `module` = "TIME_SERIES". Add a reference line ("target") at 0.5. Convert it from a line plot to a scatterplot.
4. Create a landings table.

Now, use your own data.

5. Convert your own dataset using `convert_output()`.
6. Create a spawning biomass plot. Change the theme by applying the `ggplot2::theme_classic()`.

:::

::: {.callout-caution collapse="true"}
## **Answers** for Practice exercises: Creating `stockplotr` figures & tables

For the following, use `stockplotr::example_data` as the input data.

1. Create an abundance-at-age plot. Export it as an rda to your working directory.

**Answer**:

```{r eval=FALSE}
stockplotr::plot_abundance_at_age(
  dat = stockplotr::example_data,
  make_rda = TRUE,
  figures_dir = getwd()
)
```

2. Create a recruitment time series plot. Change the `unit_label` and `scale_amount` so that the y axis shows recruitment in thousands of fish. Set `module` = "TIME_SERIES".

**Answer**:

```{r eval=FALSE}
stockplotr::plot_recruitment(
  dat = stockplotr::example_data,
  unit_label = "fish",
  scale_amount = 1000,
  module = "TIME_SERIES"
)
```

3. Create a fishing mortality plot. Ensure it does not group by fleet. Set `module` = "TIME_SERIES". Add a reference line ("target") at 0.5. Convert it from a line plot to a scatterplot.

**Answer**:

```{r eval=FALSE}
stockplotr::plot_fishing_mortality(
  dat = stockplotr::example_data,
  group = "none",
  module = "TIME_SERIES",
  ref_line = c("target" = 0.5),
  geom = "point"
)
```

4. Create a landings table.

**Answer**:

```{r eval=FALSE}
stockplotr::table_landings(
  dat = stockplotr::example_data
)
```

Now, use your own data.

5. Convert your own dataset using `convert_output()`.

**Answer**:

```{r eval=FALSE}
convert_output(
  file = "path/to/your_report_file"
  # Specify these arguments if you wish: model, fleet_names,   save_dir
)
```

6. Create a spawning biomass plot. Change the theme by applying the `ggplot2::theme_classic()`.

**Answer**:

```{r eval=FALSE}
stockplotr::plot_spawning_biomass(
  dat = your_converted_data_object
) +
  ggplot2::theme_classic()
```

:::