::: {.callout-tip collapse="true"}

# Adaptation of NEFSC Template to asar

Dan Hennen, the NEFSC representative on the workflows steering committee, has designed a sleek workflow to help aid the transition from the typical management track assessment to NOAA standard guidelines. The standard guidelines executive summary mimics closely the current management track reports, so here we will automate the executive summary in this section of the report.

Use the following code to download files into your working directory for the next process:

```{r}
#| eval: false
get_nefsc_files <- function(dir){
  NEFSCtoASAR_folder <- file.path(dir, "ExportLegacyNEFSCtoASARproject")
  testStocks_folder <- file.path(NEFSCtoASAR_folder, "testStocks")
  asarreport_folder <- file.path(NEFSCtoASAR_folder, "ASARreportFiles")
  # if (!file.exists(dir)) {
  dir.create(NEFSCtoASAR_folder)
  dir.create(testStocks_folder)
  dir.create(asarreport_folder)
  # }
  file_names_to_download <- c(
    # "ASARreportFiles",
    "CheckLatexInstall.R",
    "create_asar_object.R",
    "Example.R",
    "MapAutoUpdateToAsar.R",
    "plot_survey_indices.R",
    "plot_total_catch.R",
    "table_brp1.1.R",
    "table_catch_status1.1.R",
    "table_projections1.1.R",
    "testAdditionalStocks.R",
    "TestMoreStocksReportOnly.R"
  )
  teststocks_files <- c(
    "BSBUNITAutoAss.RData",
    "BUTUNITAutoAss.RData",
    "CODGBAutoAss.RData",
    "CODWGOMAutoAss.RData",
    "SCUNITAutoAss.RData" 
  )
  asarreport_files <- c(
    "01_executive_summary.qmd",
    "in-header.tex",
    "preamble.R" 
  )
  
  for (i in file_names_to_download) {
     cli::cli_alert_info("📥 Downloading {i}...")
    download.file(
      glue::glue("https://raw.githubusercontent.com/nmfs-ost/workflows-workshop/nefsc-transition/resources/ExportLegacyNEFSCtoASARproject/{i}"), 
      glue::glue("{NEFSCtoASAR_folder}/{i}"),
      mode = "wb"
      )
  }
  
  for (i in teststocks_files) {
     cli::cli_alert_info("📥 Downloading {i}...")
    download.file(
      glue::glue("https://raw.githubusercontent.com/nmfs-ost/workflows-workshop/nefsc-transition/resources/ExportLegacyNEFSCtoASARproject/testStocks/{i}"), 
      glue::glue("{testStocks_folder}/{i}"),
      mode = "wb"
      )
  }
  
  for (i in asarreport_files) {
     cli::cli_alert_info("📥 Downloading {i}...")
    download.file(
      glue::glue("https://raw.githubusercontent.com/nmfs-ost/workflows-workshop/nefsc-transition/resources/ExportLegacyNEFSCtoASARproject/ASARreportFiles/{i}"),
      glue::glue("{asarreport_folder}/{i}"),
      mode = "wb"
      )
  }
  
  message("✅ Download complete.")
}
get_nefsc_files(getwd())
```

The following steps will guide you through how to adapt the asar workflow and template and migrate the current content in your documents to the new ones:

  1. Change paths in TestMoreStocksReportOnly.R and source it.
  
  2. Copy the 3 files in ASARreportFiles and paste them into the report directory made by step 1 (and overwrite). 
  
```{r}
#| eval: false
file.copy(
  from = list.files(file.path(getwd(), "ExportLegacyNEFSCtoASARproject", "ASARreportFiles"), full.names = TRUE),
  to = file.path(getwd(), "report"),
  recursive = FALSE,
  overwrite = TRUE
)
```

  3. Move in-header.tex to report/support_files and overwrite (this just adds package float)
  
```{r}
#| eval: false
fs::file_move(
  file.path(getwd(), "report", "in-header.tex"),
  file.path(getwd(), "report", "support_files", "in-header.tex")
)
```

  4. change paths at the top of report/preamble.R

  5. Add 
  
```
execute:
  echo: false
  message: false
  warning: false
```
      to the yaml in the skeleton.qmd

  6. Delete lines 66 - 71 
    # load converted output from stockplotr::convert_output()
    load(fname)
    # Call reference points and quantities below
    output <- out_new |>
      dplyr::mutate(estimate = as.numeric(estimate),
        uncertainty = as.numeric(uncertainty)) 
        
    from the skeleton and *render*.

:::
