
::: {.callout-caution collapse="true"}
## Practice exercises: `create_template()`

1. Run `create_template()`. Fill in your own values for `office`, `region`, `species`, `spp_latin`, and `year`. Use default values for all remaining arguments.

2. Whoops! You forgot to add the report author's name and affiliation (yours!). Open the report skeleton and edit the yaml with this information.

3. You realize that your report will frequently reference another species: *Patrickicus starrus*. Add a new parameter to the report's yaml so that you can reference "Patrickicus starrus" with a simple "ps".

4. Open up the Executive Summary qmd file and add a sentence that references your new parameter.

:::

::: {.callout-caution collapse="true"}
## **Answers** for Practice exercises: `create_template()`

1. Run `create_template()`. Fill in your own values for `office`, `region`, `species`, `spp_latin`, and `year`. Use default values for all remaining arguments.

**Answer**: Run code similar to the following, adding in your own values for the arguments below.

```{r eval = FALSE}
create_template(
    office = "SEFSC",
    region = "Southeast Coast",
    species = "Atlantic Bigeye Tuna",
    spp_latin = "Thunnus obesus",
    year = 2026
)
```

2. Whoops! You forgot to add the report author's name and affiliation (yours!). Open the report skeleton and edit the yaml with this information.

**Answer**: Open the .qmd file with "skeleton" in the filename. Edit the yaml's "author" section to go from something like the first chunk to the second chunk.

```{r eval = FALSE}
# Original yaml snippet (located just beneath the title)
author:
  - name: 'FIRST LAST'
    affiliations: 
      - name: 'NOAA Fisheries' 
        address: 'ADDRESS' 
        city: 'CITY' 
        state: 'STATE' 
        postal-code: 'POSTAL CODE' 
```

```{r eval = FALSE}
# Updated yaml snippet
author:
  - name: 'Patrick Star'
    affiliations:
      - name: 'NOAA Fisheries Southeast Fisheries Science Center'
        address: '75 Virginia Beach Drive'
        city: 'Miami, FL'
        state: 'FL'
        postal-code: '33149'
```

3. You realize that your report will frequently reference another species: *Patrickicus starrus*. Add a new parameter to the report's yaml so that you can reference "Patrickicus starrus" with a simple "ps".

**Answer**: Open the .qmd file with "skeleton" in the filename. Edit the yaml's "params" section and add the following line:

```{r eval = FALSE}
# Updated yaml snippet
params:
   office: 'SEFSC'
   species: 'Atlantic Bigeye Tuna'
   spp_latin: 'Thunnus obesus'
   region: 'Southeast Coast'
   ps: 'Patrickicus starrus' # <-- new line
```

4. Open up the Executive Summary qmd file and add a sentence that references your new parameter.

**Answer**: Open the 01_executive_summary.qmd file and add a line that references the param 'ps', like the following:

```{r eval = FALSE}
"The species `r ps` was abundant throughout the study area."
```

*The sentence will render as such:*

"The species Patrickicus starrus was abundant throughout the study area."

:::